Local Sensitivity Analysis of a Supercritical Extraction Model

### 2025-02-13 Article_Sensitivity_Analsysis-Chamomile.tex
13.02

	% ===================================================
	% ABSTRACT
	% ===================================================
	\begin{abstract}
		This study investigates the process of chamomile oil extraction from chamomile flowers. A parameter-distributed model, consisting of a set of partial differential equations, was used to describe the governing mass transfer phenomena between solid and fluid phases under supercritical conditions using carbon dioxide as the solvent. The concept of quasi-one-dimensional flow was applied to reduce the number of spatial dimensions. The flow of carbon dioxide is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The physical properties of the solvent were estimated using the Peng-Robinson equation of state. Laboratory experiments were conducted under various, but constant operating conditions of $30 - 40~^\circ C$, $100 - 200$ bar and $3.33-6.67 \cdot 10^{-5}$ kg/s. The local sensitivity analysis method was applied to evaluate the robustness of the model parameters by investigating the impact of infinitesimally small changes in the model parameters and controls on the model outputs. This study focuses on analysing the effect of pressure on the model state space and extraction yield. {\color{blue}It was found that the model is the most sensitive to the controls if operated close to the critical point.}

@@

	The applications of supercritical carbon dioxide are not limited only to an extraction process but can also be used for impregnation, as described by \citet{Weidner2018} or \citet{Machado2022}. Impregnation is defined as modifying the properties of bulk substances by physically or chemically binding/adsorbing impregnates to a bulk material or surface, such as the hydrophobization of surfaces. The main advantage of using supercritical CO$_2$ is that, after depressurization, it desorbs from the surface and evaporates, leaving a solvent-free product. On the other hand, the main disadvantage of using carbon dioxide for impregnation is the low solubility of many drugs of interest.
	{\color{blue}The study of \citet{Ameri2020} investigates the loading of lansoprazole into polymers using supercritical carbon dioxide and examines how various parameters, such as temperature, pressure, and time, affect the drug loading efficiency. The results indicate that increasing any of these parameters enhances drug loading, with temperature having the most significant impact.	The study of \citet{Fathi2022} explores the use of supercritical carbon dioxide to enhance the bioavailability of ketoconazole by impregnating it into water-soluble polymers, specifically polyvinylpyrrolidone and hydroxypropyl methylcellulose. Utilizing a Box-Behnken design, the researchers optimized the impregnation process by varying pressure (150–250 bar), temperature (308–328 K), and time (60–180 minutes), achieving increase in drug loadings ranging.}

	Another application of supercritical CO$_2$ is nanoparticle formation, as investigated by \citet{Padrela2018}, \citet{Franco2021} or \citet{Sodeifian2022}. Supercritical carbon-dioxide-assisted technologies enable the production of different morphologies of different sizes, including nanoparticles and nanocrystals, by modulating the operating conditions. Supercritical fluid-based processes have advantages over techniques conventionally employed to produce nanosized particles or crystals, such as reduced use of toxic solvents. Moreover, the CO$_2$ is removed from the final product by simple depressurization. 
	{\color{blue} \citet{Sodeifian2018} investigates the solubility of Letrozole, a poorly water-soluble anticancer drug, in supercritical carbon dioxide with and without menthol as a solid co-solvent. The addition of menthol increased Letrozole's solubility by 7.1 times compared to supercritical CO$_2$ alone. Using the rapid expansion of supercritical solutions with solid co-solvent method, the average particle size of Letrozole was reduced to the nanoscale. Temperature was found to have the most significant impact on nanoparticle size reduction, while pressure had the least effect. 
	The study of \citet{SaadatiArdestani2020} explores the preparation of phthalocyanine green nano pigment using supercritical carbon dioxide as an antisolvent. The researchers employed the gas antisolvent technique to achieve nano-sized particles of the pigment. This method offers a potential approach for producing nano pigments with enhanced properties. The findings contribute to the development of advanced materials in the field of pigments and coatings.
	The study of \citet{Sodeifian2019} investigates the production of amiodarone hydrochloride nanoparticles using an ultrasonic-assisted rapid expansion of supercritical solution into a liquid solvent method. By optimizing parameters such as pressure, temperature, and polymeric stabilizer concentration, the researchers achieved significant particle size reduction. Characterization techniques confirmed the successful formation of nanoparticles with improved properties.
	}

@@

	{\color{blue}Depending on the needs one of two approaches can be considered while developing a mathematical model. A model, which is based on a multiple regression can be used if the relation between inputs and outputs can be simplified. \citet{Sodeifian2017a} investigated an influence of pressure, temperature and particle size on the extraction efficiency of oil from dracocephalum kotschyi boiss seed. The second order polynomial model was applied to obtain the corresponding response surface and to identify the optimum operating conditions. 
	The study of \citet{Sodeifian2017b} investigates the of essential oil from Eryngium billardieri, focusing on optimizing extraction conditions and developing a mathematical model based on the f the second order polynomial to predict the process yield. The researchers employed a simulated annealing algorithm to optimize parameters such as pressure, temperature, and extraction time, aiming to maximize oil efficiency.

	Alternatively, a more complex model can be derived and applied to cover not only the input-output relations, but the phenomena occurring in the system. This approach allows for more detailed representation the system behaviour, but requires better understanding of the mass transfer and more rigorous experiments.}

@@

	{\color{blue}The primary aim of this study is to analyse the influence of changes in operating conditions on the process model developed by \citet{Sliczniuk2024}. Understanding how variations in parameters and control variables affect the model's behaviour is essential for improving its accuracy, robustness, and practical applicability. To achieve this, a comprehensive sensitivity analysis is employed, serving as a systematic approach to assess the impact of model parameters and controls on the outputs.

	The results of this sensitivity analysis are valuable for multiple purposes. First, they help identify sources of uncertainty within the model, highlighting which parameters most significantly influence predictions. Second, they provide insights for potential model simplification, allowing the exclusion of non-essential parameters without compromising accuracy. Finally, the analysis can detect inconsistencies or errors by uncovering unexpected relationships between inputs and outputs, which may signal flaws in the model's assumptions or structure. }

@@

	{\color{blue}The paper of \citet{Maly1996} addresses the development of numerical methods and software tools for sensitivity analysis in differential-algebraic equation systems. The paper discusses the mathematical foundations of the local sensitivity analysis methods employed, providing a thorough analysis of their performance. The authors presented variety of problems and to prove the efficient of their method.

	In his \citet{Turanyi1990} presents a comprehensive review of sensitivity analysis methods applied to spatially homogeneous, constant-parameter chemical kinetic systems. The work delves into various sensitivity analysis techniques, including concentration sensitivity, rate sensitivity, and feature sensitivity analysis. The paper explores practical applications of these methods in chemical kinetics, demonstrating their utility in understanding and predicting system behavior.

	\citet{Fishtik1997} applied the local sensitivity analysis of an multiple chemical reaction system with several pure condensed phases and one ideal gas phase. It was shown that the system's response may be presented as a sum of contributions coming only from response equilibria defined by gas phase species. The effect of temperature, pressure and initial amounts of species is analyzed for gasification of coal.

	\citet{Saltelli2005} provide a comprehensive review of sensitivity analysis methods applicable to chemical modeling. The authors discuss both local and global techniques, emphasizing the importance of global sensitivity analysis in understanding complex chemical systems. The manuscript highlight the role of variance-based methods, in quantifying the influence of input parameters on model outputs.

	Unlike previous studies, this work focuses on derivative-based local sensitivity analysis to evaluate the robustness and predictive accuracy of a mathematical model for chamomile oil extraction under varying operational conditions. The analysis specifically examines the impact of operating condition changes, providing new insights into model development and optimizing extraction performance.}

@@

	where $c_f$ represents the solute concentration in the fluid phase, $r_e$ is the mass transfer kinetic term and $D^M_e$ is the {\color{blue}axial dispersion coefficient}.

@@

	{\color{blue}The choice between the two formulations should consider the system's properties and the underlying mass transfer assumptions. In systems where external mass transfer dominates, the resistance to solute transport primarily occurs at the fluid-solid interface, typically governed by the diffusion of solute across the fluid boundary layer. Under these conditions, the Bulley formulation is more appropriate, as it explicitly emphasizes the fluid-phase concentration ($c_f$) as the driving force for mass transfer through the term $c_f - c_p^*(c_s)$. Conversely, when internal mass transfer dominates, the resistance to solute transport is controlled by diffusion or adsorption processes within the solid phase. In such cases, the Reverchon formulation is more suitable, as it accounts for deviations in the equilibrium through the term $c_s - c_s^*(c_f)$.}

@@

	The method of lines is used to transform the process model equations into a set of ODEs denoted by $G(x;\Theta)$. The backward finite difference is used to approximate the first-order derivative, while the central difference scheme approximates the second-order derivative $z$ direction. The length of the fixed bed is divided into $N_z$, i.e. equally distributed points in the $z$ direction. The state-space model after discretization is denoted by $x$ and defined as follows:

	{\footnotesize
		\begin{align*} \label{discretization}
			\dot{x} &= \cfrac{d x}{d t} = 
			\begin{bmatrix}
				\cfrac{d c_{f,1}}{d t} 	  \\
				\vdots					  \\
				\cfrac{d c_{f,N_z}}{d t} \\
				\\ \hline \\
				\cfrac{d c_{s,1}}{d t} 	  \\
				\vdots					  \\
				\cfrac{d c_{s,N_z}}{d t} \\
				\\ \hline \\
				\cfrac{d h_1}{d t} 	  \\
				\vdots 					  \\
				\cfrac{d h_{N_z}}{d t} \\
				\\ \hline \\
				\cfrac{d P}{d t} \\
				\\ \hline \\
				\cfrac{d y}{d t}
			\end{bmatrix}
			=
			\underbrace{\begin{bmatrix}
					G_1 \left( c_f,c_s,h; \Theta \right)\\ 
					\vdots\\ 
					G_{N_z} \left( c_f,c_s,h; \Theta \right)\\ 
					\\ \hline \\ \\
					G_{N_z+1} \left( c_f,c_s,h; \Theta \right)\\ 
					\vdots\\
					G_{2N_z} \left( c_f,c_s,h; \Theta \right)\\ 
					\\ \\ \hline \\ 
					G_{2N_z+1} \left( c_f,c_s,h; \Theta \right) \\
					\vdots\\
					G_{3N_z} \left( c_f,c_s,h; \Theta \right) \\ 
					\\ \\ \hline \\
					G_{3N_z+1} \left( c_f,c_s,h; \Theta \right) \\
					\\ \\ \hline \\
					G_{3N_z+2} \left( c_f,c_s,h; \Theta \right) 
			\end{bmatrix}}_{G \left( x; \Theta \right),} 
	\end{align*} }
	where ${\color{black}x} \in \mathbb{R}^{N_x = 3N_z+2} $ and $\Theta \in \mathbb{R}^{N_\Theta =  N_{\theta} + N_u } $, $N_{\theta}$ is the number of parameters and $N_{u}$ is the number of control variables.

	For a derivative to be conservative, it must form a telescoping series. In other words, only the boundary terms should remain after adding all terms coming from the discretization over a grid, and the artificial interior points should be cancelled out. Discretization is applied to the conservative form of the model to ensure mass conservation.

@@

	{\color{blue}One disadvantage of using local sensitivity coefficients is that their reflect the absolute changes, which might be difficult to compare among other models or scenarios. The normalized local sensitivity coefficients, defined as $\tilde{S}(x;\Theta) = \frac{d G(x;\Theta)}{d\Theta} \frac{\Theta}{G(x;\Theta)}$, which can be interpreted as the percentage change in the output per percentage change in the parameter, making it a unitless. Although, the interpretation is straightforward it should be noted that extrapolation for large change of the parameter might leads to incorrect results. In the case of dynamic systems, the outcome of local sensitivity analysis quantifies how the state of the system or its output varies over time with respect to small perturbations in parameters. If the perturbation is too large, then the definition of a derivative (or sensitivity equations) does not hold unless the system is linear. Consequently, the conclusions can not be scaled up to any size of parameter perturbation. Such a property of the local sensitivity analysis, makes it difficult to recreate in a laboratory conditions due to finite precision of machines.

	Figure \ref{fig:Sensitivty_P_y_norm} illustrates the time evolution of the normalized sensitivity coefficients of the extraction yield. Unlike the raw sensitivities shown in Figure \ref{fig:Sensitivty_P_y}, the normalized sensitivities do not start from zero. This is due to the absence of the exact initial time point on the logarithmic time axis. The pressure change impacts the system instantaneously and uniformly, leading to a rapid initial change in the normalized sensitivities. This behavior is analogous to a step change, where the system's response to the perturbation is immediate. Initially, the sensitivity curves remain flat. The increase in normalized sensitivity begins when the solute reaches the system's outlet and becomes detectable. After reaching their respective maxima, the sensitivity curves decrease as the concentration gradient driving the extraction diminishes over time. Ultimately, all curves approach zero, reflecting that all the solute was removed from the solid phase.}

	\begin{figure}[!ht]
		\centering
		\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip,width=\columnwidth]{/Results_sensitivity/Yield_Multiple_normalized.png}
		\caption{The effect of $P_{in}$ change on $y(t)$ after normalization.}
		\label{fig:Sensitivty_P_y_norm}
	\end{figure}

@@

	Although local sensitivity analysis can be performed with respect to any model parameter, this study focuses specifically on the effect of pressure. Under the selected operating conditions, it was observed for all the cases that a pressure increment enhances mass transfer, leading to a faster mass transfer of solute from the particles. Consequently, negative sensitivities can be observed in the diagram for the solid phase. The corresponding response in the fluid phase is characterized by positive-valued sensitivities, indicating that a greater amount of solute is transported into the fluid phase. As a result, the extraction yield is improved and characterized by positive sensitivities. Given that the results of the local sensitivity analysis depend on the operating point, the analysis was repeated at different pressures.

	{\color{blue}The normalized sensitivity coefficients provide a basis for generalizing the discussion of the results and comparing the influence of various parameters. The normalized sensitivity results were found to vary within the range of -0.17 to 0.6. The overall effect of each parameter can be evaluated by integrating the sensitivity curves. Similar to the analysis of yield sensitivity, the integration was performed using an extended simulation until all sensitivities approached zero. The cumulative integral values at pressures of 100, 125, 150, 175, and 200 bars are 31487, 8548, 4462, 2825, and 1939, respectively. The positive sign of these values indicates that increased pressure enhances the extraction yield under all investigated operating conditions. The order of magnitude of the integrals allows to identify at which operating condtions the pressure increase is the most influential.}

>>>>

	% ===================================================
	% ABSTRACT
	% ===================================================
	\begin{abstract}
	This study investigates the process of chamomile oil extraction from chamomile flowers. A parameter-distributed model, consisting of a set of partial differential equations, was used to describe the governing mass transfer phenomena between solid and fluid phases under supercritical conditions using carbon dioxide as the solvent. The concept of quasi-one-dimensional flow was applied to reduce the number of spatial dimensions. The flow of carbon dioxide is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. The physical properties of the solvent were estimated using the Peng-Robinson equation of state. Laboratory experiments were conducted under various, but constant operating conditions of $30 - 40~^\circ C$, $100 - 200$ bar and $3.33-6.67 \cdot 10^{-5}$ kg/s. The local sensitivity analysis method was applied to evaluate the robustness of the model parameters by investigating the impact of infinitesimally small changes in the model parameters and controls on the model outputs. This study focuses on analysing the effect of pressure on the model state space and extraction yield. It was found that the model is the most sensitive to the controls if operated close to the critical point.

@@

	The applications of supercritical carbon dioxide are not limited only to an extraction process but can also be used for impregnation, as described by \citet{Weidner2018} or \citet{Machado2022}. Impregnation is defined as modifying the properties of bulk substances by physically or chemically binding/adsorbing impregnates to a bulk material or surface, such as the hydrophobization of surfaces. The main advantage of using supercritical CO$_2$ is that, after depressurization, it desorbs from the surface and evaporates, leaving a solvent-free product. On the other hand, the main disadvantage of using carbon dioxide for impregnation is the low solubility of many drugs of interest.
	The study of \citet{Ameri2020} investigates the loading of lansoprazole into polymers using supercritical carbon dioxide and examines how various parameters, such as temperature, pressure, and time, affect the drug loading efficiency. The results indicate that increasing any of these parameters enhances drug loading, with temperature having the most significant impact. \citet{Fathi2022} explored the use of supercritical carbon dioxide to enhance the bioavailability of ketoconazole by impregnating it into water-soluble polymers, specifically polyvinylpyrrolidone and hydroxypropyl methylcellulose. Utilizing a Box-Behnken design, the researchers optimized the impregnation process by varying pressure, temperature, and time, achieving increase in drug loadings ranging.

	Another application of supercritical CO$_2$ is nanoparticle formation, as investigated by \citet{Padrela2018}, \citet{Franco2021} or \citet{Sodeifian2022}. Supercritical carbon-dioxide-assisted technologies enable the production of different morphologies of different sizes, including nanoparticles and nanocrystals, by modulating the operating conditions. Supercritical fluid-based processes have advantages over techniques conventionally employed to produce nanosized particles or crystals, such as reduced use of toxic solvents. Moreover, the CO$_2$ is removed from the final product by simple depressurization. 
	\citet{Sodeifian2018} investigated the solubility of Letrozole, a poorly water-soluble anticancer drug, in supercritical carbon dioxide with and without menthol as a solid co-solvent. The addition of menthol increased Letrozole's solubility by 7.1 times compared to supercritical CO$_2$ alone. Using the rapid expansion of supercritical solutions with solid co-solvent method, the average particle size of Letrozole was reduced to the nanoscale. Temperature was found to have the most significant impact on nanoparticle size reduction, while pressure had the least effect. 
	The study of \citet{SaadatiArdestani2020} explored the preparation of phthalocyanine green nano pigment using supercritical carbon dioxide as an antisolvent. The researchers employed the gas antisolvent technique to achieve nano-sized particles of the pigment.
	\citet{Sodeifian2019} analysed the production of amiodarone hydrochloride nanoparticles using an ultrasonic-assisted rapid expansion of supercritical solution into a liquid solvent method. By optimizing parameters such as pressure, temperature, and polymeric stabilizer concentration, the researchers achieved significant particle size reduction. Characterization techniques confirmed the successful formation of nanoparticles with improved properties.

@@

	Depending on the needs one of two approaches can be considered while developing a mathematical model for the extraction process. A model, which is based on a multiple regression can be used if the relation between inputs and outputs is the only of interests. \citet{Sodeifian2017a} investigated an influence of pressure, temperature and particle size on the extraction efficiency of oil from dracocephalum kotschyi boiss seed. The second order polynomial model was applied to obtain the corresponding response surface and to identify the optimum operating conditions. 
	The study of \citet{Sodeifian2017b} investigates the of essential oil from Eryngium billardieri, focusing on optimizing extraction conditions and developing a mathematical model based on the second order polynomial to predict the process yield. The researchers employed a simulated annealing algorithm to optimize parameters such as pressure, temperature, and extraction time, aiming to maximize oil efficiency.

	Alternatively, a first principle model can be derived and applied to cover not only the input-output relations, but the phenomena occurring in the system. This approach allows for more detailed representation the system behaviour, but requires deeper understanding of the underlying physics and more rigorous experiments.

@@

	The primary aim of this study is to analyse the influence of changes in operating conditions on the process model developed by \citet{Sliczniuk2024}. Understanding how variations in parameters and control variables affect the model's behaviour is essential for improving its accuracy, robustness, and practical applicability. To achieve this, a comprehensive sensitivity analysis is employed, serving as a systematic approach to assess the impact of model parameters and controls on the outputs. The results of this sensitivity analysis are valuable for multiple purposes. First, they help identify sources of uncertainty within the model, highlighting which parameters most significantly influence predictions. Second, they provide insights for potential model simplification, allowing the exclusion of non-essential parameters without compromising accuracy. Finally, the analysis can detect inconsistencies or errors by uncovering unexpected relationships between inputs and outputs, which may signal flaws in the model's assumptions or structure.

@@

	The study by \citet{Maly1996} addressed the development of numerical methods and software tools for sensitivity analysis in differential-algebraic equation systems. It thoroughly examines the mathematical foundations of the local sensitivity analysis methods employed, providing an in-depth analysis of their performance. The authors presented a variety of problems to demonstrate the efficiency and applicability of their approach.

	In a his work, \citet{Turanyi1990} presented a comprehensive review of sensitivity analysis methods for spatially homogeneous, constant-parameter chemical kinetic systems. This study explored a range of sensitivity analysis techniques, including concentration sensitivity, rate sensitivity, and feature sensitivity analysis. Practical applications of these methods in chemical kinetics were highlighted, showcasing their utility in understanding and predicting system behavior.

	\citet{Fishtik1997} applied local sensitivity analysis to a complex chemical reaction system involving multiple pure condensed phases and one ideal gas phase. The study demonstrated that the system's response could be expressed as a sum of contributions from equilibrium states defined by gas-phase species. The impact of temperature, pressure, and initial species amounts was specifically analyzed in the context of coal gasification.

	In their comprehensive review, \citet{Saltelli2005} examined sensitivity analysis methods applicable to chemical reaction modeling. They discussed both local and global techniques, emphasizing the role of global sensitivity analysis in understanding complex chemical systems. The authors highlighted the importance of variance-based methods in quantifying the influence of input parameters on model outputs by discussing results of a few simulation analysis.

	Unlike previous studies, this work focuses on derivative-based local sensitivity analysis to evaluate the robustness and predictive accuracy of a mathematical model for chamomile oil extraction under varying operational conditions. The analysis specifically investigates the effects of changes in operating conditions, offering novel insights into model development and the optimization of extraction performance.

@@

	where $c_f$ represents the solute concentration in the fluid phase, $r_e$ is the mass transfer kinetic term and $D^M_e$ is the axial dispersion coefficient.

@@

	The choice between the two formulations should consider the system's properties and the underlying mass transfer assumptions. In systems where external mass transfer dominates, the resistance to solute transport primarily occurs at the fluid-solid interface, typically governed by the diffusion of solute across the fluid boundary layer. Under these conditions, the Bulley formulation is more appropriate, as it explicitly emphasizes the fluid-phase concentration ($c_f$) as the driving force for mass transfer through the term $c_f - c_p^*(c_s)$. Conversely, when internal mass transfer dominates, the resistance to solute transport is controlled by diffusion or adsorption processes within the solid phase. In such cases, the Reverchon formulation is more suitable, as it accounts for deviations in the equilibrium through the term $c_s - c_s^*(c_f)$.

@@

	\begin{figure}[!h]
		\centering
		{\footnotesize
			\begin{align*}
				\dot{x} &= \cfrac{d x}{d t} = 
				\begin{bmatrix}
					\cfrac{d c_{f,1}}{d t} 	  \\
					\vdots					  \\
					\cfrac{d c_{f,N_z}}{d t} \\
					\\ \hline \\
					\cfrac{d c_{s,1}}{d t} 	  \\
					\vdots					  \\
					\cfrac{d c_{s,N_z}}{d t} \\
					\\ \hline \\
					\cfrac{d h_1}{d t} 	  \\
					\vdots 					  \\
					\cfrac{d h_{N_z}}{d t} \\
					\\ \hline \\
					\cfrac{d P}{d t} \\
					\\ \hline \\
					\cfrac{d y}{d t}
				\end{bmatrix}
				=
				\underbrace{\begin{bmatrix}
						G_1 \left( c_f,c_s,h; \Theta \right)\\ 
						\vdots\\ 
						G_{N_z} \left( c_f,c_s,h; \Theta \right)\\ 
						\\ \hline \\ \\
						G_{N_z+1} \left( c_f,c_s,h; \Theta \right)\\ 
						\vdots\\
						G_{2N_z} \left( c_f,c_s,h; \Theta \right)\\ 
						\\ \\ \hline \\ 
						G_{2N_z+1} \left( c_f,c_s,h; \Theta \right) \\
						\vdots\\
						G_{3N_z} \left( c_f,c_s,h; \Theta \right) \\ 
						\\ \\ \hline \\
						G_{3N_z+1} \left( c_f,c_s,h; \Theta \right) \\
						\\ \\ \hline \\
						G_{3N_z+2} \left( c_f,c_s,h; \Theta \right) 
				\end{bmatrix}}_{G \left( x; \Theta \right)} 
		\end{align*} }
	\caption{Discretized state-space}
	\label{fig:discretization}
	\end{figure}

	The method of lines is used to transform the process model equations into a set of ODEs denoted by $G(x;\Theta)$. For a derivative to be conservative, it must form a telescoping series. In other words, only the boundary terms should remain after adding all terms coming from the discretization over a grid, and the artificial interior points should be cancelled out. Discretization is applied to the conservative form of the model to ensure mass conservation. The backward finite difference is used to approximate the first-order derivative, while the central difference scheme approximates the second-order derivative $z$ direction. The length of the fixed bed is divided into $N_z$, i.e. equally distributed points in the $z$ direction. The state-space model after discretization is denoted by $x$ and defined as follows the Figure \ref{fig:discretization}, where ${\color{black}x} \in \mathbb{R}^{N_x = 3N_z+2} $ and $\Theta \in \mathbb{R}^{N_\Theta =  N_{\theta} + N_u } $, $N_{\theta}$ is the number of parameters and $N_{u}$ is the number of control variables.

@@

	\begin{figure}[!hb]
		\centering
		\includegraphics[trim = 0.0cm 0.0cm 0.0cm 0.0cm,clip,width=\columnwidth]{/Results_sensitivity/Yield_Multiple_normalized.png}
		\caption{The effect of $P_{in}$ change on $y(t)$ after normalization.}
		\label{fig:Sensitivty_P_y_norm}
	\end{figure}

	One disadvantage of using local sensitivity coefficients is that they reflect absolute changes, which can be challenging to compare across different models or scenarios. To address this, normalized local sensitivity coefficients (defined as: $\tilde{S}(x;\Theta) = \frac{d G(x;\Theta)}{d\Theta} \frac{\Theta}{G(x;\Theta)}$) are used. They can be interpreted as the percentage change in the output per percentage change in the parameter. Their dimensionless nature makes the invariant under rescaling of model parameters.

	Figure \ref{fig:Sensitivty_P_y_norm} illustrates the time evolution of the normalized sensitivity coefficients of the extraction yield. Unlike the raw sensitivities shown in Figure \ref{fig:Sensitivty_P_y}, the normalized sensitivities do not start from zero. This is due to the absence of the exact initial time point on the logarithmic time axis. The pressure change impacts the system instantaneously and uniformly, leading to a rapid initial change in the normalized sensitivities. This behavior is analogous to a step change, where the system's response to the perturbation is immediate. Initially, the sensitivity curves remain flat. The increase in normalized sensitivity begins when the solute reaches the system's outlet and becomes detectable. After reaching their respective maxima, the sensitivity curves decrease as the concentration gradient driving the extraction diminishes over time. Ultimately, all curves approach zero, reflecting that all the solute was removed from the solid phase.

	While this interpretation of normalized sensitivities is straightforward, it is important to note that extrapolating for large parameter changes can lead to inaccurate results.In dynamic systems, local sensitivity analysis quantifies how the system state or output varies over time in response to small parameter perturbations. If the perturbation is large, the derivative-based definition of sensitivity (or sensitivity equations) becomes invalid unless the system is linear. Consequently, the conclusions from local sensitivity analysis cannot be reliably scaled to large parameter changes. This limitation makes reproducing of sensitivity results in laboratory conditions challenging, given the finite precision of experimental equipment.

@@

	Although local sensitivity analysis can be performed with respect to any model parameter, this study focuses specifically on the effect of pressure. Under the selected operating conditions, it was observed for all the cases that a pressure increment enhances mass transfer, leading to a faster mass transfer of solute from the particles. Consequently, negative sensitivities can be observed in the diagram for the solid phase. The corresponding response in the fluid phase is characterized by positive-valued sensitivities, indicating that a greater amount of solute is transported into the fluid phase. As a result, the extraction yield is improved and characterized by positive sensitivities. Given that the results of the local sensitivity analysis depend on the operating point, the analysis was repeated at different pressures. The normalized sensitivity coefficients provide a basis for generalizing the discussion of the results and comparing the influence of various parameters. The normalized sensitivity results were found to vary within the range of -0.17 to 0.6. The overall effect of each parameter can be evaluated by integrating the sensitivity curves. Similar to the analysis of yield sensitivity, the integration was performed over extended simulation time (until all sensitivities approached zero). The cumulative integral values at pressures of 100, 125, 150, 175, and 200 bars are 31487, 8548, 4462, 2825, and 1939, respectively. The positive sign of these values indicates that increased pressure enhances the extraction yield under all investigated operating conditions. The order of magnitude of the integrals allows to identify at which operating condtions the pressure increase is the most influential.
